---
title: "Time tRavelling: Introduction to Longitudinal models in R"
author: "Vivek Jason"
---

### Tutorial 1: Exploring Correlation Structures in GEE

Objective: In this exercise, we'll fit three different correlation structures to our data: unstructured, exchangeable, and AR1 (first-order autoregressive). By comparing their performance, you'll gain insight into how different correlation structures impact your GEE model results.

**Step 1**: Fit the Unstructured Model: Start by fitting the model using an unstructured correlation structure as you've done before.

**Step 2**: Fit the Exchangeable Model: Use an exchangeable correlation structure and refit the model.

**Step 3**: Fit the AR1 Model: Now, fit the model using the first-order autoregressive structure.

**Step 4**: Compare AIC Values: For each model, extract the AIC (Akaike Information Criterion) value, which serves as a measure of model goodness-of-fit.

**Step 5**: Analyse: Compare the AIC values. Lower AIC values indicate better model fits. Based on this criterion, which correlation structure seems the most suitable for our data?

Solution:

```{r}

```

Now, based on the AIC values, which correlation structure would you consider the most appropriate for this dataset?

### Tutorial 2: Exploring mixed models

In the dataset we've been working with, let's explore how the relationship between `sedentary bebehavioursed)` and `depression (dep)` varies across different clusters (represented by the variable `sdmvstra`). To do this, fit four different mixed models:

1.  Fixed intercept, fixed slope
2.  Fixed intercept, random slope
3.  Random intercept, fixed slope
4.  Random intercept, random slope

Step 1: Fixed Intercept, Fixed Slope: This is essentially a linear regression, as neither the intercept nor the slope is allowed to vary across clusters. Step 2: Fixed Intercept, Random Slope: The model has a constant intercept but allows the slope for `sed` to vary across clusters. Step 3: Random Intercept, Fixed Slope: The intercept can change across clusters, but the relationship (slope) between `sed` and `dep` remains constant. Step 4: Random Intercept, Random Slope: Both the intercept and the slope for `sed` can vary across clusters.

```{r}

```

By fitting and comparing these models, you can better understand the relationship between sedentary behavior and depression and how this relationship might differ across clusters or groups.
